---
title: "almond_informal_sensitivity"
author: "Kyle Burt, Michelle Moreno"
format: html
editor: visual
---

# Libraries

```{r}
library(tidyverse)
library(purrr)
library(ggpubr)
library(here)
```

# Load data

```{r}
clim <- read.table(here("data", "clim_04202025.txt"), 
                   header = TRUE, sep = " ")
```

# Data management

### Filter data

```{r}
filtered_clim <- clim |>
  filter(month %in% c(1, 2)) |>
  select(-wy, -tmax_c)
```

### Create data for total precipitation calculation

```{r}
jan_precip_summary <- filtered_clim |>
  filter(month == 1) |>
  group_by(year) |>
  summarise(jan_precip = sum(precip, na.rm = TRUE), 
            .groups = "drop")
```

### Create data for average minimum temperature calculation

```{r}
feb_TminC_summary <- filtered_clim |>
  filter(month == 2) |>
  group_by(year) |>
  summarise(feb_tmin_c = mean(tmin_c, na.rm = TRUE), 
            .groups = "drop")
```

### Join the two summaries

```{r}
clim_inputs <- left_join(feb_TminC_summary, jan_precip_summary, by = "year")
```

# Calculate almond yield anomaly for each year

```{r}
source("R/combined_functions.R")

yield_profit_summary <- compute_profit_fromyield(
  min_temp = clim_inputs$feb_tmin_c,
  P = clim_inputs$jan_precip,
  year = clim_inputs$year
)
```

# Informal sensitivity analysis using temperature and precipitation

### Define ranges for sensitivity: +/-20%

```{r}
min_temp_range <- seq(0.8, 1.2, by = 0.1)

precip_range <- seq(0.8, 1.2, by = 0.1)
```

### Expand grid of combinations

```{r}
sensitivity_grid <- expand.grid(
  temp_factor = min_temp_range,
  precip_factor = precip_range
)
```

### Compute scenarios

```{r}
sensitivity_results <- sensitivity_grid %>%
  mutate(results = pmap(list(temp_factor, precip_factor), function(tf, pf) {
    compute_profit_fromyield(
      min_temp = clim_inputs$feb_tmin_c * tf,
      P = clim_inputs$jan_precip * pf,
      year = clim_inputs$year
    )
  })) %>%
  unnest(results) %>%
  # Add temp_factor and precip_factor back in after unnesting
  mutate(temp_factor = rep(sensitivity_grid$temp_factor, each = nrow(clim_inputs)),
         precip_factor = rep(sensitivity_grid$precip_factor, each = nrow(clim_inputs)))
```

# Plot of yield anomaly by year with parameter uncertainty

```{r}
ggplot(sensitivity_results, aes(x = year, y = yield_anomaly, 
                                group = interaction(temp_factor, 
                                                    precip_factor))) +
  geom_line(alpha = 0.3) +
  stat_summary(fun = mean, geom = "line", color = "blue", size = 1.2) +
  labs(title = "Yield Anomaly by Year under Parameter Uncertainty",
    y = "Yield Anomaly",
    x = "Year") +
  theme_bw()
```

# Plot of yield anomaly sensitivity to parameters (mean across years)

### Calculate mean yield anomaly

```{r}
summary_sensitivity <- sensitivity_results |>
  group_by(temp_factor, precip_factor) |>
  summarise(mean_yield_anomaly = mean(yield_anomaly), .groups = "drop")
```

### Graph

```{r}
ggplot(summary_sensitivity, aes(x = temp_factor, y = precip_factor, 
                                fill = mean_yield_anomaly)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma") +
  labs(title = "Sensitivity of Yield Anomaly to Temperature and Precipitation", x = "Temperature Factor", y = "Precipitation Factor", 
       fill = "Mean Yield Anomaly") +
  theme_bw()
```
